---
# title: "Neural Network Example in R"
output: 
  github_document:
    html_preview: false
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

### Goal
The goal is to create a working example of a basic deep neural network in R.
This example uses the tidyverse style of R. The main goal is clarity and explaining the concept, not speed or practical application.

### Packages
```{r library, message = FALSE, warning = FALSE}
library(tidyverse)
```


### Training data
The training data used in this example is fake data. And this example starts of with an output that is lineairly related to two input variables.
```{r fake_data, message = FALSE, warning = FALSE}
data <- tibble(
  input_value_1 = c(0:10) / 10,
  input_value_2 = c(10:0) / 50,
  output_value = c(0:10) / -5 + 1
)

data |> head(7)
```


### Setup neural network
```{r nn_parameters, message = FALSE, warning = FALSE}
input_neurons  <- 2
output_neurons <- 1
layers         <- 2
layer_neurons  <- 3
```

Every layer, except for the input layer, has a weigth matrix and bias vector associated with it. These are also related to the computational steps when passing the signal forward. So the structure to store these parameters will look like this:

```{r nn_datastructure, echo = FALSE}
list(
  first_step = list(
    weigths = matrix(1, nrow = 2, ncol = 2),
    biases = rep(1, 2)
  ),
  second_step = list(
    weigths = matrix(1, nrow = 2, ncol = 2),
    biases = rep(1, 2)
  )
)
```

First a matrix of all weights (w) from the input neurons to the first layer of hidden neurons is created. This is an input_neurons x layer_neurons matrix, where input_neurons is the amount of neurons in the input layer, and layer_neurons is the amount of neurons per hidden layer. And this is repeated until the last layer to the output neurons is innitialised.

```{r nn_w_and_b, message = FALSE, warning = FALSE}
neurons_per_layer <- c(input_neurons, rep(layer_neurons, layers), output_neurons)

create_weights_and_biases <- function(neurons_from, neurons_to) {
  list(
    weights = matrix(runif(neurons_from * neurons_to, -1, 1), nrow = neurons_from),
    biases = runif(neurons_to, -1, 1)
  )
}

weights_and_biases <-
  map2(neurons_per_layer[-length(neurons_per_layer)],
       neurons_per_layer[-1],
       create_weights_and_biases)

weights_and_biases

```



# Forward propagation
